---
title: "plot WRF: Elevation"
author: "Sergio Ibarra-Espinosa"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plot WRF: Elevation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we attempt to provide useful code to generate figures from WRF outputs based on known galleries. For instance, NCL and WRF-Python provides extensive examples for plotting WRF outputs. Therefore, we aim to replicate some of these. Our approach to read wrfout files is based on eixport which relies r packages with GDAL bindings such as raster and stars. We do not try to provide a full gallery, instead, some basics and necessary plots to inspire other R used and receive more examples so share with the community.

## packages

- [eixport](https://atmoschem.github.io/eixport/) read and manipulate wrf files.
- [raster](https://github.com/rspatial/raster) for gridded and raster data.
- [stars](https://github.com/r-spatial/stars) for gridded and raster data.
- [cptcity](https://github.com/ibarraespinosa/cptcity) colour palettes.
- [sf](https://github.com/r-spatial/sf) for spatial vector data.

```{r}
library(eixport)
library(raster)
library(stars)
library(cptcity)
library(sf)
```

Based on [NCL](https://www.ncl.ucar.edu/Applications/Scripts/wrf_gsn_1.ncl):

```{r}
wrfo <- "/media/sergio/ext41/wrfo_sebrazil/wrfout_d02_2014-11-03_000000"
```

Reading HGT from wrfout
```{r}
HGT <- wrf_get(wrfo,  "HGT", as_raster = T)
HGT <- HGT[[1]] # by default one variable with each time, so we select one
```

Adding coastlines and cropping for our study area
```{r}
library(rnaturalearth)
cl <- ne_countries(scale = "small", returnclass = "sf")
cl <- st_transform(cl, 31983)
HGT <- st_transform(st_as_stars(HGT), 31983)
cl <- st_cast(st_crop(cl, HGT), "LINESTRING")
```

Find colour palette for elevation
```{r}
find_cpt("elevation")
```


```{r, fig.width=6,fig.height=4}
#spplot(HGT, main = "HGT using spplot", scales=list(draw = TRUE),
#       col.regions = cpt("grass_elevation"),
#       sp.layout = list("sp.lines", as_Spatial(cl), col = "black"))
```   

```{r, fig.width=6,fig.height=5}
cols <- classInt::classIntervals(HGT$HGT_2014.11.03_00.00.00, n = 100, style = "pretty")
plot(HGT, axes = T, main = "HGT using stars", col = cpt("grass_elevation", n = length(cols$brks) - 1), breaks = cols$brks, reset = F)
plot(cl$geometry, add= T, col = "black")
```
