---
title: "WRF out statistics"
author: "Sergio Ibarra-Espinosa"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WRF out statistics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we attempt to provide useful code to generate figures from WRF outputs based on known galleries. For instance, NCL and WRF-Python provides extensive examples for plotting WRF outputs. Therefore, we aim to replicate some of these. Our approach to read wrfout files is based on eixport which relies r packages with GDAL bindings such as raster and stars. We do not try to provide a full gallery, instead, some basics and necessary plots to inspire other R used and receive more examples so share with the community.

## packages

- [eixport](https://atmoschem.github.io/eixport/) read and manipulate wrf files.
- [raster](https://github.com/rspatial/raster) for gridded and raster data.
- [stars](https://github.com/r-spatial/stars) for gridded and raster data.
- [cptcity](https://github.com/ibarraespinosa/cptcity) colour palettes.
- [sf](https://github.com/r-spatial/sf) for spatial vector data.

```{r}
library(eixport)
library(raster)
library(stars)
library(cptcity)
library(sf)
```

Based on [NCL](https://www.ncl.ucar.edu/Applications/Scripts/wrf_gsn_1.ncl):

First lets get a summary of a WRF output file
```{r}
wrfo <- "/home/sergio/R/x86_64-pc-linux-gnu-library/4.3/helios/extras/wrfout_d01_2020-01-01_01%3A00%3A00_sub.nc"
(dt <- wrf_meta(wrfo)$vars)
```
Now we can select some variables

```{r}
(vars <- dt$vars)
```

Now get some statistics

```{r}
wrf_summary(wrfo, vars = c("T2"))
```

