---
title: "plot WRF: Elevation"
author: "Sergio Ibarra-Espinosa"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plot WRF: Elevation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we attempt to provide useful code to generate figures from WRF outputs based on known galleries. For instance, NCL and WRF-Python provides extensive examples for plotting WRF outputs. Therefore, we aim to replicate some of these. Our approach to read wrfout files is based on eixport which relies r packages with GDAL bindings such as raster and stars. We do not try to provide a full gallery, instead, some basics and necessary plots to inspire other R used and receive more examples so share with the community.

## packages

- [eixport](https://atmoschem.github.io/eixport/) read and manipulate wrf files.
- [raster](https://github.com/rspatial/raster) for gridded and raster data.
- [stars](https://github.com/r-spatial/stars) for gridded and raster data.
- [cptcity](https://github.com/ibarraespinosa/cptcity) colour palettes.
- [sf](https://github.com/r-spatial/sf) for spatial vector data.

```{r}
library(eixport)
library(raster)
library(stars)
library(cptcity)
library(sf)
```

Based on [NCL](https://www.ncl.ucar.edu/Applications/Scripts/wrf_gsn_1.ncl):

Reading HGT from wrfout
```{r}
HGT <- wrf_get("/media/sergio/ext4/wrfoETA_COM_VEIN/wrfout_d03_2014-10-03_00:00:00", 
               "HGT", as_raster = T)
HGT <- HGT[[1]] # bythefault one variable with each time, so we select one
```

Adding coastlines and cropping for our study area
```{r}
cl <- st_read("/media/sergio/ext4/coasts_shapefile/gshhg-shp-2.3.7/GSHHS_shp/f/GSHHS_f_L1.shp")
cl <- st_cast(st_crop(cl, st_as_stars(HGT)), "LINESTRING")
```

Find colour palette for elevation
```{r}
find_cpt("elevation")
```


```{r, fig.width=6,fig.height=4}
plot(HGT, main = "HGT using plot", col = cpt("grass_elevation"))
plot(cl, add= T, col = "black")
```

```{r, fig.width=6,fig.height=4}
spplot(HGT, main = "HGT using spplot", scales=list(draw = TRUE),
       col.regions = cpt("grass_elevation"),
       sp.layout = list("sp.lines", as_Spatial(cl), col = "black"))
```   

```{r, fig.width=6,fig.height=5}
# hgt <- st_as_stars(HGT)
# plot(hgt, axes = T, main = "HGT using stars", col = cpt("grass_elevation"), breaks = "fisher", reset = F)
# plot(cl$geometry, add= T, col = "black")
```
